Think deep for reef resilience

### 2025-08-10 paper/paper_draft.qmd
Minor adjustments to paper and 3_ script to ensure smooth run

For the purpose of scenario assessment, the resulting timeseries for each model run were
scaled by the total area of the reef to represent cover as a percentage of the total
estimated reef area. The proportion of each functional group occupying the reef (group
evenness) was also investigated as a metric for timeseries analysis, but was found to not
sufficiently capture the declined condition of reefs suffering from very low absolute coral
cover **\[TI: Ben, is the sentence here really important to include?\]**. These timeseries
of coral cover were then clustered based on their temporal behaviour.

@@

```{julia}
#| echo: false
#| warning: false
#| message: false
#| results: asis
#| label: tbl-cluster-cid
#| tbl-cap: Complexity invariant distance (CID) between low and high cluster median timeseries in each management area under each GCM. CID values are calculated for percentage cover timeseries (@fig-gcm-cover-trajectories) and Degree Heating Week (DHW) timeseries (@fig-gcm-dhw-trajectories). Complexity invariant distance represents the euclidean distance between timeseries, corrected by the complexity of the two timeseries. CID values are used to quantify overall differences between these two cluster levels at management area and GCM levels. CID values for percentage cover and DHW timeseries are not directly comparable as the timeseries use different units.

@@

using PrettyTables, Markdown
cd("../src")

include("../src/common.jl")

cluster_regression = CSV.read(joinpath(output_path, "cluster_regression_outputs/ACCESS-ESM1-5_bioregion_cluster_regression_coefficients.csv"), DataFrame)
cluster_regression.Name = replace.(cluster_regression.Name, ["ACCESS-ESM1-5_bioregion_cluster_cats: " => "Cluster "])

for name in names(cluster_regression)
    if isa(cluster_regression[:, name], Vector{Float64})
        cluster_regression[:, name] = round.(cluster_regression[:, name]; sigdigits=2)
    end
end

header = ["Response level ~ Term", "β estimate", "SE", "t-value", "Pr > |t|", "2.5%", "97.5%"]
pretty_table(io, cluster_regression, backend=Val(:markdown), header=header)

@@

Our findings do not suggest that reefs in ‘high’ cover clusters have greater thermal
tolerance than reefs in lower coral cover clusters. This is in contrast with literature
suggesting corals have the capacity to adapt to increasing climate change pressures
[@humanes2022; @lachs2024], including the investigation by Bairos-Novak et al.
[-@bairos-novak2021] that revealed high heritability of bleaching and survival response
traits, used to inform thermal tolerance trait distributions in this investigation. The lack
of influence of thermal tolerance on reef outcomes is likely due to reefs being impacted by
high levels of thermal stress early in projected conditions and being unable to build
thermal tolerance to DHW levels greater than 9°C-weeks. **\[TI note: This paragraph needs to
be revisited\]**

It is a common conclusion in coral reef ecology that a primary facet of reef resilience is
the degree of larval connectivity between reefs and the supply of recruits that a reef
receives following disturbance [@hock2017; @mumby2008; @torres2024]. In contrast, our
results suggest that there are few differences between resilient and non-resilient reefs in
connectivity metrics such as the ratio of outgoing to incoming larval connectivity and total
connectivity strength. Furthermore, our results suggest no relationship between connectivity
and the ability of reefs to maintain positive carbonate budgets under future conditions.
There has been contrasting evidence [@bode2018; @hock2017] surrounding whether recent
generation connectivity modelling is able to identify reefs that are important for recovery
processes, with findings dependent on the larval connectivity model used. These
investigations were performed using prior generations of larval connectivity models
parameterised for 2008-2013 [@hock2017] and 1996-2002 [@bode2018] time periods. Using
updated 1km resolution larval connectivity across the whole GBR, with a limited temporal
sample of particle flows, our results suggest that larval connectivity alone is not as
important for reef resilience and recovery as environmental refugia such as depth.

@@

![Depth cutoff for reefs](images/reef_depth_included_cutoff.png){#sfig-depth-cutoff
fig-align="center" width="14.8cm" height="9.5cm"}

@@

![![](images/paste-1.png)Heatmaps indicating reef cluster assignments within each bioregion
across the five GCMs. Green indicates a low cover cluster assignment, orange represents a
medium cover assignment and blue represents assignment of a reef to a high coral cover
cluster. The vertical width of reef elements varies depending on the number of reefs within
the bioregion (displayed in the y axis). The x axis represents the different GCM levels
(EC-Earth3-Veg, ACCESS-ESM1-5, ACCESS-CM2, NorESM2-MM and
GFDL-CM4).](../outputs/ADRIA_results/HighResCoralStress/figs/GCM_bioregion_cluster_assignments.png){#sfig-gcm-bioregion-cluster-assignment
fig-align="center" width="14.82cm"}

>>>>

For the purpose of scenario assessment, the resulting timeseries for each model run were
scaled by the total area of the reef to represent cover as a percentage of the total
estimated reef area. These timeseries
of coral cover were then clustered based on their temporal behaviour.

@@

```{julia}
#| echo: false
#| warning: false
#| message: false
#| results: asis
#| label: tbl-cluster-cid
#| tbl-cap: Complexity invariant distance (CID) between low and high cluster median timeseries in each management area under each GCM. CID values are calculated for percentage cover timeseries (@fig-gcm-cover-trajectories) and Degree Heating Week (DHW) timeseries (@sfig-gcm-dhw-trajectories). Complexity invariant distance represents the euclidean distance between timeseries, corrected by the complexity of the two timeseries. CID values are used to quantify overall differences between these two cluster levels at management area and GCM levels. CID values for percentage cover and DHW timeseries are not directly comparable as the timeseries use different units.

@@

# using PrettyTables, Markdown
# cd("../src")

# include("../src/common.jl")

# cluster_regression = CSV.read(joinpath(output_path, "cluster_regression_outputs/ACCESS-ESM1-5_bioregion_cluster_regression_coefficients.csv"), DataFrame)
# cluster_regression.Name = replace.(cluster_regression.Name, ["ACCESS-ESM1-5_bioregion_cluster_cats: " => "Cluster "])

# for name in names(cluster_regression)
#     if isa(cluster_regression[:, name], Vector{Float64})
#         cluster_regression[:, name] = round.(cluster_regression[:, name]; sigdigits=2)
#     end
# end

# io = IOBuffer()

# header = ["Response level ~ Term", "β estimate", "SE", "t-value", "Pr > |t|", "2.5%", "97.5%"]
# pretty_table(io, cluster_regression, backend=Val(:markdown), header=header)

# md_string = (String(take!(io)))
# Markdown.parse(md_string)
```

@@

Our findings suggest that reefs in ‘high’ cover clusters do not have greater thermal
tolerance than reefs in lower coral cover clusters. This is in contrast with literature
suggesting corals have the capacity to adapt to increasing climate change pressures
[@humanes2022; @lachs2024], including the investigation by Bairos-Novak et al.
[-@bairos-novak2021] that revealed high heritability of bleaching and survival response
traits, used to inform thermal tolerance trait distributions in this investigation. The lack
of influence of thermal tolerance on reef outcomes is likely due to reefs being impacted by
high levels of thermal stress early in projected timeseries and being unable to build
thermal tolerance to DHW levels greater than 9°C-weeks. **\[TI note: This paragraph needs to
be revisited\]**

It is a common conclusion in coral reef ecology that a primary facet of reef resilience is
the degree of larval connectivity between reefs and the supply of recruits that a reef
receives following disturbance [@hock2017; @mumby2008; @torres2024]. In contrast, our
results suggest that there are few differences between resilient and non-resilient reefs in
connectivity metrics such as the ratio of outgoing to incoming larval connectivity and total
connectivity strength. Furthermore, our results suggest no relationship between connectivity
and the ability of reefs to maintain positive carbonate budgets under future conditions.
There has been contrasting evidence [@bode2018; @hock2017] surrounding whether recent
generation connectivity modelling is able to identify reefs that are important for recovery
processes, with findings dependent on the larval connectivity model used. These
investigations were performed using prior generations of larval connectivity models
parameterised for 2008-2013 [@hock2017] and 1996-2002 [@bode2018] time periods. Using
updated 1km^2^ resolution larval connectivity across the whole GBR, with a limited temporal
sample of particle flows, our results suggest that larval connectivity alone is not as
important for reef resilience and recovery as environmental refugia such as depth.

@@

![Depth cutoff for reefs](../outputs/ADRIA_results/HighResCoralStress/figs/reef_depth_included_cutoff.png){#sfig-depth-cutoff
fig-align="center" width="14.8cm" height="9.5cm"}

@@

![Heatmaps indicating reef cluster assignments within each bioregion
across the five GCMs. Green indicates a low cover cluster assignment, orange represents a
medium cover assignment and blue represents assignment of a reef to a high coral cover
cluster. The vertical width of reef elements varies depending on the number of reefs within
the bioregion (displayed in the y axis). The x axis represents the different GCM levels
(EC-Earth3-Veg, ACCESS-ESM1-5, ACCESS-CM2, NorESM2-MM and
GFDL-CM4).](../outputs/ADRIA_results/HighResCoralStress/figs/GCM_bioregion_cluster_assignments.png){#sfig-gcm-bioregion-cluster-assignment
fig-align="center" width="14.82cm"}
